---
title: "Homework_02: Data wrangling"
author: "Mustafa Kamal Shahadat"
date: "2023-03-18"
output: html_document
---

#### Required packages
```{r}
library(tidyverse)
library(janitor)
```
#### Load data from directory
```{r}
getwd()
arsenic <- read.csv("bd_arsenic_data_raw.csv")
head(arsenic)
```


### Tasks
1. Use janitor::row_to_names() remove text from column heading
2. Remove empty rows and columns using janitor::remove_empty()
3. Clean column names using janitor::clean_names()
4. Use as.numeric and as.factor arguments in dplyr::mutate_at() function to convert ‘chr’ columns to numeric and factors accordingly
5. Find duplicate records and remove them
6. Run all above functions with pipe.


```{r}
## Create a data folder
dataFolder<-"L:\\Training\\Digital Soil Mapping Training (BARC 17-18 Jan 2023)\\DSM_On-line Training 27 Feb to 13 June 2023\\Data\\Bangladesh\\"

# Loading data from data folder
df<-read_csv(paste0(dataFolder,"bd_arsenic_data_raw.csv")) %>%
  glimpse()
#head(df)
```


#### 1.	Use janitor::row_to_names() remove text from column heading
```{r}
df.01 = df %>% 
  janitor::row_to_names(1, remove_row = TRUE, remove_rows_above = TRUE) %>%
  glimpse()
```

#### 2.	Remove empty rows and columns using janitor::remove_empty()
```{r}
df.02 = df.01 %>% 
  janitor::remove_empty() %>%
  glimpse()
```
#### 3.	Clean column names using janitor::clean_names()
```{r}
df.03 = df.02 %>%
  janitor::clean_names() %>%
  glimpse()
```
```{r}
colnames(df.03)
```

#### 4.	Use as.numeric and as.factor arguments in dplyr::mutate_at() function to convert ‘chr’ columns to numeric and factors accordingly
```{r}
df.04= df.03 %>%
     dplyr::mutate_at(5:19, as.numeric) %>%
     dplyr::mutate_at(2:4, as.factor) %>%
     glimpse()
```
#### 5.	Find duplicate records and remove them
```{r}
df.04 %>% janitor::get_dupes(id)
```

```{r}
#  check the duplicate records in the this data
df.05= df.04 %>% 
     dplyr::distinct(id,.keep_all = TRUE) %>%
     glimpse()
```




#### 6.	Run all above functions with pipe.
```{r}
df_clean = df %>% 
  janitor::row_to_names(1, remove_row = TRUE, remove_rows_above = TRUE) %>%
  janitor::remove_empty() %>%
  janitor::clean_names() %>%
  dplyr::mutate_at(5:19, as.numeric) %>%
  dplyr::mutate_at(2:4, as.factor) %>%
  dplyr::distinct(id,.keep_all = TRUE) %>%
     glimpse()
```

```{r}
# save clean data
write_csv(df_clean,"bd_arsenic_data_final.csv" )
```



















