---
title: "Homework_03"
author: "Mustafa Kamal Shahadat"
date: "2023-03-18"
output: html_document
---
### Load required packages
```{r}
library(plyr) # Note: load 'plyr' before 'dplyr'
library(dplyr)
library(tidyverse)
library(car)
library(carData)
library(nortest)
library(outliers)
library(report)
library(RColorBrewer)
library(ggstatsplot)
library(gtsummary)
library(ggExtra)
library(gridExtra)
library(xtable)
library(flextable)
```
```{r}
#install.packages("tidyverse", dependencies = TRUE)
#install.packages("car", dependencies = TRUE)
#install.packages("nortest", dependencies = TRUE)
#install.packages("outliers", dependencies = TRUE)
#install.packages("plyr", dependencies = TRUE)
#install.packages("report", dependencies = TRUE)
#install.packages("RColorBrewer", dependencies = TRUE)
#install.packages("ggstatsplot", dependencies = TRUE)
#install.packages("gtsummary", dependencies = TRUE)
#install.packages("ggExtra", dependencies = TRUE)
#install.packages("gridExtra", dependencies = TRUE)
#install.packages("xtable", dependencies = TRUE)
#install.packages("flextable", dependencies = TRUE)
```

```{r}
# Load data files
df<- read_csv("bd_arsenic_data.csv")
df.gh<- read_csv("bd_green_house.csv")
```
### Tasks
1. Check the data structure with str() function
```{r}
str(df)
```
```{r}
str(df.gh)
```

2. Check missing values if any.
```{r}
sum(is.na(df))
sum(is.na(df.gh))
```

3. Create a summary statistics table with these variables: GAs, WAs, SAs, SPAs, SAoAs, SAoFe from arsenic data.
```{r}
summary(df$GAs) 
```
```{r}
summary(df$WAs)

```
```{r}
summary(df$SAs)

```
```{r}
summary(df$SPAs)

```
```{r}
summary(df$SAoAs)

```
```{r}
summary(df$SAoFe)
```

4. Create a summary statistics by group (Landtype) using above mention variables.
```{r}
summary_stat <- df %>% 
  # First select  numerical columns
  dplyr::select(GAs, WAs, SAs, SPAs, SAoAs, SAoFe) %>%
  # get summary statistics
  dplyr::summarise_all(funs(Min = round(min(.),2), 
                      Q25 = round(quantile(., 0.25),2), 
                      Median = round(median(.),2), 
                      Q75 = round(quantile(., 0.75),2), 
                      Max = round(max(.),2),
                      Mean = round(mean(.),2), 
                      SD = round(sd(.),3))) %>% 
                # create a nice looking table
                tidyr::pivot_longer(everything(), 
                            names_sep = "_", 
                            names_to = c( "variable", ".value"))

# Load library
library(flextable)
# Create a table
flextable::flextable(summary_stat, theme_fun = theme_booktabs)
```

5. Create a histogram, density plot QQ-plot of Soil As (SAs) with ggplot2 package.
```{r}
ggplot(df, aes(SAs)) +
  geom_histogram()+
  # X-axis title
  xlab("SAs") + 
  # y-axis title
  ylab("Frequency")+
  # plot title
  ggtitle("Histogram of Soil SAs")
```
```{r}
ggplot(df, aes(SAs)) +
  geom_density()+
  # x-axis title
  xlab("SAs") + 
  # y-axis title
  ylab("Density")+
  # plot title
  ggtitle("Kernel Density of Soil SAs")+
    theme(
    # Center the plot title
    plot.title = element_text(hjust = 0.5))
```
```{r}
ggplot(df, aes(sample = SAs)) +
  stat_qq() + 
  stat_qq_line() +
  # x-axis title
  xlab("Theoretical Quantiles)") + 
  # y-axis title
  ylab("Sample Quantiles")+
  # plot title
  ggtitle("Q-Q plot of Soil SOC from a Normal Distribution")+
    theme(
    # Center the plot title
    plot.title = element_text(hjust = 0.5),
    # x-axis title font size
    axis.title.x = element_text(size = 14),
    # y-axis title font size
    axis.title.y = element_text(size = 14),)
```

6. Do normality test of WAs, SAs, GAs and measure the Skewness and Kurtosis of these variables.
```{r}
shapiro.test(df$WAs)
```
```{r}
shapiro.test(df$SAs)
```
```{r}
shapiro.test(df$GAs)
```
```{r}
library(moments)
moments::skewness(df$WAs)
moments::kurtosis(df$WAs)
```
```{r}
moments::skewness(df$SAs)
moments::kurtosis(df$SAs)
```
```{r}
moments::skewness(df$GAs)
moments::kurtosis(df$GAs)
```

7. Create two scatter plots with ggplot2: i) WAs vs GAs and ii) SAs vs GAs, and show them side by side (p1+p2)
```{r}
p1<- ggplot(df, aes(x=WAs, y=GAs)) +
  geom_point(size=2) +
  # add plot title
  ggtitle("Scatterplot of Soil WAs vs GAs") +
   theme(
    # center the plot title
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "gray"),
    # axis title font size
    axis.title.x = element_text(size = 14), 
    # X and  axis font size
    axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5),
    axis.text.x = element_text(size=12))+
     xlab("WAs") + 
     ylab("Soil GAs")
p1
```
```{r}
p2<- ggplot(df, aes(x=SAs, y=GAs)) +
  geom_point(size=2) +
  # add plot title
  ggtitle("Scatterplot of Soil SAs vs GAs") +
   theme(
    # center the plot title
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "red"),
    # axis title font size
    axis.title.x = element_text(size = 14), 
    # X and  axis font size
    axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5),
    axis.text.x = element_text(size=12))+
     xlab("SAs") + 
     ylab("GAs")
p2
```
```{r}
p1+p2
```

8. Create a 3D bubble plot
```{r}
#conflicts_prefer(plyr::mutate)
#conflicts_prefer(dplyr::arrange)
df %>%
  arrange(desc(SOC)) %>%
  mutate(Land_type = factor(Land_type)) %>%
  ggplot(aes(x=ELEV, y=SOC, size = SP, color=Land_type)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.2, 4), name="SP")+
     guides(color = guide_legend(title = "Land_type"))+
  ggtitle("Bubble Plot of Soil SOC vs ELEV") +
   theme(
    # center the plot title
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "gray"),
    # axis title font size
    axis.title.x = element_text(size = 14), 
    # X and  axis font size
    axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5),
    axis.text.x = element_text(size=12))+
      # add legend tittle
  guides(color = guide_legend(title = "Land Type")) + 
  xlab("ELEV") + 
  ylab("Soil OC (mg/g)") 
```

9. Create a correlation matrix table showing significant levels. Use GAs, StAs, WAs, SAs, SPAs, SAoAs, ans SAoFe variables.
```{r}
#library(Hmisc)
# create a data frame for correlation analysis
df.cor<-df %>% dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs, SAoFe) 
# correlation matrix
cor.mat<-Hmisc::rcorr(as.matrix(df.cor, type="pearson"))
cor.mat
```
```{r}
# install.packages("xtable")
library(xtable)

cor_table <- function(x){ 
  require(Hmisc) 
  x <- as.matrix(x) 
  R <- rcorr(x)$r 
  p <- rcorr(x)$P 
  
  ## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
  
  ## trunctuate the matrix that holds the correlations to three decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), 3))[,-1] 
  
  ## build a new matrix that includes the correlations with their apropriate stars 
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
  diag(Rnew) <- paste(diag(R), " ", sep="") 
  rownames(Rnew) <- colnames(x) 
  colnames(Rnew) <- paste(colnames(x), "", sep="") 
  
  ## remove upper triangle
  Rnew <- as.matrix(Rnew)
  Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
  Rnew <- as.data.frame(Rnew) 
  
  ## remove last column and return the matrix (which is now a data frame)
  Rnew <- cbind(Rnew[1:length(Rnew)-1])
  return(Rnew) 
}
```
```{r}
cor_table(df.cor)
```

10. Create a Correlation matrix plot of above correlation analysis.
```{r}
# Load GGally library
library(GGally)
GGally::ggpairs(df.cor)
```


```{r}
# Load corrplot library
library(corrplot)
# Create plot
corrplot(cor.mat$r, 
         type="upper", 
         order="hclust", 
         cex.lab = 0.5,
         tl.col = "black",
         p.mat = cor.mat$p, 
         sig.level = 0.05, insig = "blank")
```

11. Use ggplot2 to create two barplots with error bar: i) GAs and Treatment II) GAs and varieties. And show them side by side. Use bd_green_house data
```{r}
# Standard error
SE <- function(x){
  sd(x)/sqrt(length(x))
}
```
```{r}
# Load library
#library(plyr)
# Get summary statistics
summarise_GAs1<-ddply(df.gh,~Treatment, summarise, 
                Mean= round(mean(GAs), 3),
                Median=round (median(GAs), 3),
                Min= round (min(GAs),3), 
                Max= round (max(GAs),3), 
                SD= round(sd(GAs), 3), 
                SE= round (SE(GAs), 3))
# Load library
library(flextable)
# Create a table
flextable::flextable(summarise_GAs1, theme_fun = theme_booktabs)
```

```{r}
#| warning: false
#| error: false
#| fig.width: 5
#| fig.height: 4.5
#| 
p3 = ggplot(summarise_GAs1, aes(x=Treatment, y=Mean)) + 
  geom_bar(stat="identity", position=position_dodge(),width=0.5, fill="yellow") +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.2,
   position=position_dodge(.9))+
  # add y-axis title and x-axis title leave blank
  labs(y="GAs", x = "")+ 
  # add plot title
  ggtitle("Mean ± SE of GAs")+
  coord_flip()+
  # customize plot themes
  theme(
        axis.line = element_line(colour = "gray"),
        # plot title position at center
        plot.title = element_text(hjust = 0.5),
        # axis title font size
        axis.title.x = element_text(size = 14), 
        # X and  axis font size
        axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5, colour='black'),
        axis.text.x = element_text(size=12))
p3
```

```{r}
# Get summary statistics
summarise_GAs2<-ddply(df.gh,~Variety, summarise, 
                Mean= round(mean(GAs), 3),
                Median=round (median(GAs), 3),
                Min= round (min(GAs),3), 
                Max= round (max(GAs),3), 
                SD= round(sd(GAs), 3), 
                SE= round (SE(GAs), 3))
# Load library
library(flextable)
# Create a table
flextable::flextable(summarise_GAs2, theme_fun = theme_booktabs)
```


```{r}
p4 = ggplot(summarise_GAs2, aes(x=Variety, y=Mean)) + 
  geom_bar(stat="identity", position=position_dodge(),width=0.5, fill="green") +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.2,
   position=position_dodge(.9))+
  # add y-axis title and x-axis title leave blank
  labs(y="GAs", x = "")+ 
  # add plot title
  ggtitle("Mean ± SE of GAs")+
  coord_flip()+
  # customize plot themes
  theme(
        axis.line = element_line(colour = "gray"),
        # plot title position at center
        plot.title = element_text(hjust = 0.5),
        # axis title font size
        axis.title.x = element_text(size = 14), 
        # X and  axis font size
        axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5, colour='black'),
        axis.text.x = element_text(size=12))

p4
```

```{r}
p3+p4
```

12. Do t-test and show whether there is any significant difference in As concentration in rice (GAs) between low and high As soil treatments.
```{r}
# Create two dataframes
T1 <- df.gh %>% dplyr::select(Treatment, GAs) %>%
         filter(Treatment == "High As")

T2 <- df.gh %>% dplyr::select(Treatment, GAs) %>%
         filter(Treatment == "Low As")

# Two-Sample T-test
t.test(x=T1$GAs,
       y=T2$GAs, 
       paired = TRUE,
       alternative = "greater")
```

13. Do one-way and two-way ANOVA show main effect of treatment and rice variety and their interaction on rice grain As and explain the results.
```{r}
anova.one=aov (GAs ~Treatment, data = df.gh) # analysis variance 
anova (anova.one)
```
```{r}
# Load library
# library(report)
report(anova.one)
```
Two-way ANOVA
```{r}
anova.two =aov (GAs ~Treatment + Variety+ Treatment:Variety, data = df.gh)
anova (anova.two) 
```
```{r}
# Load library
# library(report)
report(anova.two)
```


14. Do multiple comparison of two main effect.

```{r}
tukey <- TukeyHSD(anova.one)
tukey
```
```{r}
pairwise.t.test(df.gh$GAs, df.gh$Treatment,
                 p.adjust.method = "bonferroni", 
                 paired=FALSE)
```
```{r}
pairwise.t.test(df.gh$GAs, df.gh$Variety,
                 p.adjust.method = "bonferroni", 
                 paired=FALSE)
```


```{r}
# Load library
#library(ggstatsplot)

p1<-ggbetweenstats(
  data = df.gh,
  x = Treatment,
  y = GAs,
  ylab = "GAs",
  type = "parametric", # ANOVA or Kruskal-Wallis
  var.equal = TRUE, # ANOVA or Welch ANOVA
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
)+
# add plot title
ggtitle("Mian Effect of Treatment") +
   theme(
    # center the plot title
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "gray"),
    # axis title font size
    axis.title.x = element_text(size = 14), 
    # X and  axis font size
    axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5),
    axis.text.x = element_text(size=12))
print(p1)
```


```{r}
# Load library
library(ggstatsplot)

p2<-ggbetweenstats(
  data = df.gh,
  x = Variety,
  y = GAs,
  ylab = "GAs",
  type = "parametric", # ANOVA or Kruskal-Wallis
  var.equal = TRUE, # ANOVA or Welch ANOVA
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
)+
# add plot title
ggtitle("Mian Effect of Variety") +
   theme(
    # center the plot title
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "gray"),
    # axis title font size
    axis.title.x = element_text(size = 14), 
    # X and  axis font size
    axis.text.y=element_text(size=12,vjust = 0.5, hjust=0.5),
    axis.text.x = element_text(size=12))
print(p2)
```

```{r}
library(patchwork)
p1+p2
```








