---
title: "Homework_04"
author: "Mustafa Kamal Shahadat"
date: "2023-03-23"
output: html_document
---
Loading Required packages
```{r}
library(tidyverse)
library(dlookr)
library(flextable)
library(janitor)
```
Loading data file
```{r}
# mf<-read_csv("bd_arsenic_data_raw.csv")
# Load data
mf<-read.csv(paste0("bd_arsenic_data_raw.csv"), header=TRUE)
```

### Task
1. Use dlookr::diagnose() function of dlookr package to do Data Quality Diagnosis all variables

Before moving to quality checking we need to do some pre-processing the data
```{r}
#deleting first row of the data table since it contains data description
mf.01 = mf %>% 
  janitor::row_to_names(1, remove_row = TRUE, remove_rows_above = TRUE) %>% 
  glimpse()
```
```{r}
#Removing empty cells
mf.02 = mf.01 %>% 
  janitor::remove_empty() %>%
  glimpse()
```
```{r}
#Cleaning variable names
mf.03 = mf.02 %>%
  janitor::clean_names() %>%
  glimpse()
```
```{r}
# Renaming variables for convenience
mf.04 <- mf.03 %>% 
        dplyr::rename("GAs" = "grain_as", "StAs" = "straw_as", "WAs" = "w_as", "SAs" = "s_as", "SPAs" = "sp_as", "SAoAs" = "sao_as")
names(mf.04)
```
```{r}
# Looking at the data structure
str(mf.04)
```
```{r}
# Checking whether we are working with data frame
is.data.frame(mf.04)
```
```{r}
# Changing character values to numeric values
mf.04 <- mf.04%>%
      dplyr::mutate_at(c("GAs", "StAs", "WAs", 'SAs', "SPAs", "SAoAs"), as.numeric)%>%
  glimpse()
```
```{r}
# Checking data quality with dlookr
dlookr::diagnose(mf.04)%>%
  flextable() 
```

2. Use dlookr::diagnose_numeric(), diagnoses numeric such as GAs,StAs, WAs, SAs, SPAs,SAoAs and SAoFe
```{r}
# First select  numerical columns
mf.04 %>% 
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
# then diagnose them
  dlookr::diagnose_numeric()%>% 
  flextable()
```
```{r}
# Checking the variable names of the data frame
names(mf.04)
```


3. Use dlookr::diagnose_category() to diagnoses the categorical variables (Land_type and variety) of a data frame.
```{r}
mf.04 %>% 
# Select categorical variables
  dplyr::select(variety, land_type, year) %>%
# then diagnose them
  dlookr::diagnose_category()%>% 
  flextable()
```


4. Use dlookr::diagnose_outlier() to diagnosis outlier of GAs, StAs, WAs, and SAs.
```{r}
mf.04 %>% dlookr::diagnose_outlier(GAs, StAs, WAs, SAs, SPAs, SAoAs)
```


5. Use dlookr::plot_outlier() to visualizes outliers of SAs
```{r}
mf.04 %>%
  dlookr::plot_outlier(dlookr::diagnose_outlier(mf.04,SAs) %>% 
                 dplyr::filter(outliers_ratio >= 0.5) %>% 
                 dplyr::select(variables) %>% 
                 unlist())
```


6. Perform a normality test on multiple numerical data (such as GAs,StAs, WAs,SAs, SPAs,SAoAs and SAoFe) using dlookr::normality() function.
```{r}
mf.04 %>% 
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
  dlookr::normality() %>%
  # sort variables that do not follow a normal distribution in order of p_value:
  dplyr::filter(p_value <= 0.01) %>% 
  dplyr::arrange(abs(p_value)) %>%
  flextable()
```


7. Use dlookr::normality() and dplyr::group_by() check normality of rice garin As (GAs) in different rice varieties.
```{r}
mf.04 %>% 
  dplyr::group_by(variety) %>%
  dlookr::normality(GAs) %>%
  dplyr:: arrange(desc(p_value)) %>%
  flextable()
```


8. Plot histogram and QQ-plot of SAs and histogram and square root transformed SAs using dlookr::plot_normality() function
```{r}
mf.04 %>% dlookr::plot_normality(SAs)
```


9. Create a descriptive statistics table for GAs, StAs, Soil, WAs, PAs using dlookr::describe() and flextable() functions
```{r}
# First select  numerical columns
des.stata<-mf.04 %>% 
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
 # then descrive them
  dlookr::describe()
flextable(des.stata)
```


10. Use dlookr::describe() and dplyr::group_by() functions calculate descriptive testatrices of SAs, SPAs, SAoAs and SAoFe in different landtypes
```{r}
mf.04 %>%
  group_by(land_type) %>% 
  dlookr::describe(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
  flextable()
```


11. Calculates the correlation coefficient of all combinations of WAs, WFe, WP, WEc, SEc, SAs, SPAs, SAoAs and SAoFe and plot them using dlookr::correlate() and plot() functions.
```{r}
# First select  numerical columns
mf.04 %>% 
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
# then diagnose them
  dlookr::correlate()%>% 
  flextable()
```


12. Calculates the correlation coefficients in two landtypes: use dlookr::correlate() and dplyr::group_by()
```{r}
mf.04 %>% 
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
# then diagnose them
  dlookr::correlate()%>% 
  plot()
```
```{r}
mf.04 %>% 
  group_by(land_type) %>%
  dplyr::select(GAs, StAs, WAs, SAs, SPAs, SAoAs) %>%
# then diagnose them
  dlookr::correlate()%>% 
  plot()
```
























