---
title: "Homework_01"
author: "Mustafa Kamal Shahadat, Senior Scientific Officer, Bangladesh Agricultural Research Institute, On-Farm Research Division, Daulatpur, Khulna "
date: "2023-03-17"
output: html_document
---

### Load 'tidyverse' package
```{r message=TRUE, warning=TRUE}
library(tidyverse)
```

### Read data files
```{r}
rice <- read.csv("bd_rice_production.csv")
head(rice)
district <- read.csv("bd_district.csv")
head(district)
division <- read.csv("bd_division.csv")
head(division)
raj_soil <- read.csv("raj_soil_data.csv")
head(raj_soil)
bd_soil <- read.csv("bd_soil_data_raw.csv")
head(bd_soil)
```

## Tasks for Homework_01

### 1: joining objects 'rice' and 'district' using inner_join() function of 'dplyr' package 

```{r}
bd_rice = dplyr::inner_join(rice, district)
```
### 2: joining objects 'bd_rice'and 'division' using inner_join() function of 'dplyr' package 
```{r}
bd_rice = dplyr::inner_join(bd_rice, division)
```
### 3: Organize column as (DIVISION_ID, DIVISION_NAME, DISTRICT_ID, DISTRICT_NAME, SEASON, YEAR, MT using dplyr::relocate()
```{r}
bd_rice <-dplyr::relocate(bd_rice, DIVISION_PCODE, DIVISION_NAME, DISTRICT_ID, DISTRICT_NAME, SEASON, YEAR, MT,
                          .after =  DIVISION_PCODE) 
head(bd_rice)
```

### 4.	Rename “DISTRICT_ID” to “DISTRICT_PCODE” using dplyr::rename()
```{r}
bd_rice <- bd_rice %>% 
  dplyr::rename("DISTRICT_PCODE" = "DISTRICT_ID")
names(bd_rice)
```
### 5.	Run all the above code using pipe function in a single line and create a data frame name df.rice.csv
```{r}
df.rice.csv = dplyr::inner_join(rice, district) %>% 
          dplyr::inner_join(division) %>%
          dplyr::relocate(DIVISION_PCODE, DIVISION_NAME, DISTRICT_ID, DISTRICT_NAME, SEASON, YEAR, MT,
                    .after =  DIVISION_PCODE) %>%
          dplyr::rename("DISTRICT_PCODE" = "DISTRICT_ID") %>%
        glimpse()
```
### 6.	Form df.rice data that you have created before, use select() function to create a data frame only with DISTRICT_NAME, YEAR and MT
```{r}
df.dist <- df.rice.csv %>% 
           dplyr::select(DISTRICT_NAME, YEAR,  MT,) %>%
           glimpse()
```
### 7.	Filter out the data from Rajshahi division using dplyr::filter()
```{r}
df.raj<-df.rice.csv %>% 
              dplyr::filter(DIVISION_NAME == "Rajshahi")
levels(as.factor(df.raj$DISTRICT_NAME))
```
### 8.	Filter (Filtering by multiple criteria) all district in Rajshahi and Dhaka divisions using dplyr::filter()
```{r}
df.div<- df.rice.csv %>% 
            dplyr::filter(DIVISION_NAME %in%c("Rajshahi", "Dhaka")) 
levels(as.factor(df.div$DISTRICT_NAME))
```
### 9.	Use | or OR in the logical condition in filter function, select data from any of the two divisions.
```{r}
df.div.01<- df.rice.csv %>% 
                    dplyr::filter(DIVISION_NAME == "Rajshahi" | DIVISION_NAME == "Dhaka")
levels(as.factor(df.div.01$DISTRICT_NAME))
```
### 10.	Apply filter create a files for the Rajshai Division only with Bogra District.
```{r}
df.raj.bogra<-df.rice.csv %>% 
         dplyr::filter(DIVISION_NAME == "Rajshahi" & DISTRICT_NAME == "Bogra")
tail(df.raj.bogra)
```
### 11.	Select districts where rice production (MT) is greater than the country average
```{r}
mean.prod <- df.rice.csv %>% 
              dplyr::filter(MT > mean(MT, na.rm = TRUE))
levels(as.factor(mean.prod$DISTRICT_NAME))
```
### 12.	Use & to select district where MT is greater than the country average of for the year 2018-2019
```{r}
mean.prod.2018.19 <- df.rice.csv %>% 
  dplyr::filter(MT > mean(MT, na.rm = TRUE) & YEAR ==2018-19)
levels(as.factor(mean.prod.2018.19$DISTRICT_NAME))
```
### 13.	Select districts starting with “P”
```{r}
dist.p <- df.rice.csv %>% 
          dplyr::filter(grepl("^P", DISTRICT_NAME))
levels(as.factor(dist.p$DISTRICT_NAME))
```
### 14.	Use raj_soil data to calculate mean of SOM using dplyr::summarise_at() function

```{r}
#Checking column names in the data file
colnames(raj_soil)
```
```{r}
# Extracting mean of SOM from raj_soil
raj_soil %>%
  dplyr::summarise_at("SOM", mean, na.rm = TRUE)%>%
  glimpse()
```
### 15.	Use bd_soil data to calculate mean of SOM and NDVI using dplyr::summarise_at() function

```{r}
raj_soil %>%
    dplyr::summarise_at(c("SOM", "NDVI"), mean, na.rm = TRUE)
```
### 16.	Use dplyr::summarise_if() to calculate mean() to the numeric columns
```{r}
raj_soil %>%
    dplyr::summarise_if(is.numeric, mean, na.rm = TRUE)
```
### 17.	Select numerical columns (pH,SOM, DEM, NDVI, NDFI) and then apply dplyr::summarise_all()*to calculate mean of these variables
```{r}
raj_soil %>%
    # First select  numerical columns
    dplyr::select(pH, SOM, DEM, NDFI, NDVI) %>%
    # get mean of all these variables
    dplyr::summarise_all(mean, na.rm = TRUE)
```

### 18.	Create a new column (MT_1000) in df.rice dataframe dividing MT column by 1000 using mutate()
```{r}
df.rice.csv %>%
    dplyr::mutate(MT_1000 = MT / 10000) %>%
    glimpse()
```
### 19.	Apply the group_by(), summarize()** and mutate() functions with pipe to calculate mean of rice production in 1000 MT by division for the year 2018-2019
```{r}
df.rice.csv %>% 
          dplyr::filter(YEAR=="2018-19") %>%
          dplyr::group_by(DIVISION_NAME) %>%
          dplyr::summarize(Prod_MT = mean(MT)) %>%
          dplyr::mutate(Prod_1000_MT = Prod_MT / 1000)  
```

```{r}
colnames(df.rice.csv)
```


### 20.	Use df.rice data to create a wider data frame using tidyr::pivot_wider() function
```{r}
rice.wider = df.rice.csv  %>%
  dplyr::select (DIVISION_NAME, DISTRICT_NAME, SEASON, YEAR, MT) %>%
  tidyr::pivot_wider(names_from = YEAR, values_from = MT) %>%
  glimpse()
```
### 21.	Convert rice.wider data to a longer format use tidyr::pivot_longer()
```{r}
names(rice.wider)
```
```{r}
rice.longer<- rice.wider %>% 
  tidyr::pivot_longer(cols= c("2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019"),
                      names_to = "YEAR", # column need to be wider
                      values_to = "MT") # data
head(rice.longer)
```
### 22.	Clean bd_soil_data_raw.csv and save as bd_soil_data.final.csv. Follow the steps you have learned from “Cleaning a messy data” section.
```{r}
# Looking at the column names
glimpse(bd_soil)
```

```{r}
# Create a file with limited number of variables using select functions and rename them
mf.geo <- bd_soil %>%
      select(x,y,DIVISION_NAME,DISTRICT_NAME, UPAZILA_NAMAE,UNION_NAME, Soil_PH,Soil_SOM,Soil_K, Soil_Ca, MG, N, P, S, B, CU, FE, MN, ZN) %>%
      rename("PH"=Soil_PH,
             "SOM" =Soil_SOM,
             "K" = Soil_K,
             "Ca" = Soil_Ca)%>% 
  glimpse()
```
```{r}
# Filter the data where records have N.D. values 
mf.geo<- mf.geo %>% 
        filter(B != "N.D.")
```

```{r}
# convert all N.D. values to empty string
mf.geo[mf.geo=="N.D."]<- ""
```

```{r}
# Here detection limits of K, Ca, N, P, S, CU, FE, and Mn are 0.1, 0.1, 0.1, 1.0, 0.05, 0.1, 1.0 and 0.1 mg/kg, respectively. We will replace the values below detection limits by half of detection limits of these nutrients. Before that we have to remove “<” and convert the all to .
mf.geo <- mf.geo %>%
  mutate_at(c("K", "Ca", "N", "P", "S", "CU", "FE", "MN"), str_replace, "<", "")%>%
  mutate_at(c("K", "Ca", "N", "P", "S", "CU", "FE", "MN"), as.numeric) %>% glimpse()
```
```{r}
# replace values below detection limits
mf.geo["K"][mf.geo["K"] == 0.1] <- 0.05
mf.geo["Ca"][mf.geo["Ca"] == 0.1] <- 0.05
mf.geo["N"][mf.geo["N"] == 0.1] <- 0.05
mf.geo["P"][mf.geo["P"] == 1.0] <- 0.5
mf.geo["S"][mf.geo["S"] == 0.05] <- 0.025
mf.geo["CU"][mf.geo["CU"] == 0.1] <- 0.05
mf.geo["FE"][mf.geo["FE"] == 1.0] <- 0.5
mf.geo["MN"][mf.geo["MN"] == 0.1] <- 0.05
```
### Checking missing values of the data
```{r}
# counting unique, missing, and median values
K<- mf.geo %>% summarise(Length = length(K),
                 na = sum(is.na(K)),
                 Min = min(K, na.rm = TRUE),
                 Max =max(K, na.rm = TRUE))

Ca<- mf.geo %>% summarise(Length = length(Ca),
                 na = sum(is.na(Ca)),
                 Min = min(Ca, na.rm = TRUE),
                 Max =max(Ca, na.rm = TRUE))

N<- mf.geo %>% summarise(Length = length(N),
                 na = sum(is.na(N)),
                 Min = min(N, na.rm = TRUE),
                 Max =max(N, na.rm = TRUE))

P<- mf.geo %>% summarise(Length = length(P),
                 na = sum(is.na(P)),
                 Min = min(P, na.rm = TRUE),
                 Max =max(P, na.rm = TRUE))

S<- mf.geo %>% summarise(Length = length(S),
                 na = sum(is.na(S)),
                 Min = min(S, na.rm = TRUE),
                 Max =max(S, na.rm = TRUE))

CU<- mf.geo %>% summarise(Length = length(CU),
                 na = sum(is.na(CU)),
                 Min = min(CU, na.rm = TRUE),
                 Max =max(CU, na.rm = TRUE))

FE<- mf.geo %>% summarise(Length = length(FE),
                 na = sum(is.na(FE)),
                 Min = min(FE, na.rm = TRUE),
                 Max =max(FE, na.rm = TRUE))

MN<- mf.geo %>% summarise(Length = length(MN),
                 na = sum(is.na(MN)),
                 Min = min(MN, na.rm = TRUE),
                 Max =max(MN, na.rm = TRUE),
                 )

#bind the data
geo.sum<- bind_rows(K, Ca, N, P, S, CU, FE, MN)

#add.row.names
row.names(geo.sum) <- c("K", "Ca", "N", "P", "S", "CU", "FE", "MN")
```

```{r}
head(geo.sum)
```

```{r}
# Omitting missing values
newdata <- na.omit(mf.geo)
glimpse(newdata)
```



```{r}
# Data imputation with mean 
mf.geo.new<-mf.geo %>% mutate_at(vars(K, Ca, N, P, CU, FE, MN),~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)) %>% glimpse()
```

```{r}
# Checking mission values
mf.geo.new %>%summarise(sum(is.na(K)))
```
```{r}
# save clean data
write_csv(mf.geo.new,"bd_soil_data.final.csv" )
```











